---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```



### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv
```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```



### Task 3: Calculate the percentage change in each category between Fall 2024 and Fall 2021 using `mutate()`

```{r}
umd_fee_pct <- umd_fees |> mutate(pct_change = (`Fall 2024` - `Fall 2021` ) / `Fall 2021` * 100)

```


### Task 4: Describe our results

Write a few sentences about what you found in the data. What's the story here?

Fees at the University of Maryland have increased every year since 2021. The biggest increase has been in how much students pay every year for shuttle bus services. In 2021 the amount students pay for bus services is almost 40% higher than it was in 2021. This increase is what I find most notbale and what would be most newsorthy in this data.




When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
