# JOUR405: Statistics for Journalists
## Midterm Exam - Spring 2025

Name: YOUR NAME HERE
Zephan Matteson

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences. Start by loading the tidyverse and any other libraries you think you might need.

```{r}
library(tidyverse)
```

## Part 1: Restaurant Health Inspections (15 points)

You want to understand how restaurants in Montgomery County are performing on health inspections. The first dataset contains restaurant health inspection scores for restaurants in Montgomery County. The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv` and complete these tasks:

```{r}
Montgomery_Inspection_Scores <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```

### Tasks:
1. Calculate the mean and standard deviation of compliance scores (5 points)
2. Create a histogram of the compliance scores with a vertical line showing the mean (5 points)
3. Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? (5 points).

```{r}
Montgomery_Inspection_Scores |> summarize(meancompliance = mean(compliance_score), sdcompliance = sd(compliance_score))

```
```{r}
Montgomery_Inspection_Scores |> 
  ggplot() + 
  geom_histogram(aes(x = compliance_score), binwidth = 5) +
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed", size = 1) + 
   geom_vline(aes(xintercept = mean(compliance_score) - sd(compliance_score)), color = "blue", linetype = "dashed", size = 1) + 
   geom_vline(aes(xintercept = mean(compliance_score) + sd(compliance_score)), color = "blue", linetype = "dashed", size = 1) 
```
The data for compliance scores is skewed to the right meaning that the mean for this data is going to be slightly misleading. I think that the data is being pulled down by outlier bad scores. The story I then i think would be how well restraunts in general are doing with their compliance scores and inspections.

## Part 2: High School Athletics (25 points)

You are reporting a story about high school sports participation in Maryland and want to see if there are differences between boys and girls. The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and sex. Load the data from: `https://raw.githubusercontent.com/example/md_hs_sports_2024.csv` and complete these tasks:

```{r}
sport_participation <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv")
```


### Tasks:
1. Calculate the correlation between boys' and girls' participation (5 points)
2. Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. (5 points)
3. Create a scatterplot showing this relationship, adding a line of best fit (5 points)
4. In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

```{r}
sport_participation |> summarize(correlation = cor(boys, girls))
```
```{r}
sport_participation <- sport_participation |> mutate(total = (boys + girls), girls_pct = (girls / total))
```

```{r}
sport_participation |> 
  ggplot() +
  geom_point(aes(x=girls, y=boys)) + 
  geom_smooth(aes(x=girls, y=boys), method="lm")
```
The correlation between boys and girls playing sports in maryland is very high. If a lot of boys are playing sports, then a lot of girls will also be playing sports. This is shown by the coefficient, but also by observing the data points on the scatterplot and seeing how they group in a positive linear direction around a definite line.

## Part 3: Public Transit Ridership (20 points)

You are investigating public transit ridership in the Washington, D.C. area and want to understand the patterns of daily bus and rail ridership. The third dataset contains daily bus and rail ridership totals from WMATA for the past year. Load the data from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv and do the following:

```{r}
public_ridership <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")
```


### Tasks:
1. Calculate the average bus and rail ridership and standard deviation using summarize() (5 points)
2. Using the process you used in class, take a random sample daily ridership numbers and calculate the sample means and deviations for bus and rail. The number in the sample is up to you, but explain why you chose what you did. Compare this to the stats you generated in step 1. (5 points)
3. Using group_by() and summarize(), calculate the means for bus and rail ridership for each weekday. Describe the overall pattern of ridership for bus and rail - which days stand out and why? Are there differences between bus and rail in the standard deviation values? (10 points)


```{r}
public_ridership |> summarize(meanbus = mean(bus), meanrail = mean(rail), sdbus = sd(bus), sdrail = sd(rail))
```

```{r}
set.seed(1234)

sample_ridership_100 <- public_ridership |> 
  sample_n(100)

sample_ridership_100 |> summarize(meanbus100 = mean(bus), sdbus100 = sd(bus), meanrail100 = mean(rail), sdrail100 = sd(rail))
```
I decided to take a sample of 100 because that was roughly 1/4 of the population. This seemed to work pretty well as generally the mean and sd for the sample matched up relatively well to the total population. The sample of 100 jsut generally gave higher numbers but usually only by a couple thousand which in context didn't seem like a huge deal to me.

```{r}
public_ridership |> summarize(meanbus = mean(bus), meanrail = mean(rail), sdbus = sd(bus), sdrail = sd(rail))
```


## Part 4: Maryland Car Theft Rates (20 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

```{r}
Car_Theft <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```


### Tasks:
1. Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate (5 points)
2. Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median, and what percentage of all car thefts occur in those counties? (5 points)
3. Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? (10 points)


```{r}
Car_Theft <- Car_Theft |> mutate(theft_2022 = `2022` / population * 10000, theft_2023 = `2023` / population * 1000)
```

```{r}
Car_Theft |> summarise(medianrate2023 = median(theft_2023), totalteft = sum(`2023`))
```
```{r}

```

## Part 5: Data Analysis Scenario (20 points)

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences (no code!) explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)
2. What visualizations would help readers understand the trends? (5 points)
3. What additional context or data would you need to make this a complete story? (5 points)

I would first calculate what the average response time for the data this year is, and then do the same for probably the last ten years of data, and compare. If by looking at the yearly data over a period of time the response is significantly worse this year that would verify the data. Also calculating the mean response time for the first six months of the year and then the last six months of the year would also tell me whether something is happening that year. I think some sort of histogram would prbably be the best visual, or perhaps a scatterplot that shows respsone times decreasing negatively as the year goes on if the data backs that up. Again the additional context would be the data from the past years response times. Is repsonse time super variable year to year? Do response times get slower at certain points every year? Were there major cuts to the staff taht have resulted in slower times? Perhaps that would be the story instead. 

### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Before submitting, clear your environment and run the entire notebook

Good luck!
