---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: <https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv>. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```

### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

ANSWER HERE The data provided records the number of incidents there were of 18 different crimes in PG County. The most obvious one that I would look at first as a reporter is homicides, that is what people are really going to want to know about. The other big things I would want to look at would be violent crimes like assault and sex offences. Those are the big crimes that readers would care about, and then also potentially theft, and car accident numbers as well as secondary stories. The only thing that complicates it in my opinion is that the murder rate seems to be pretty steady so the biggers story is probably in a crime. That crime likely being assault or theft in the county.

### Task 4: Calculating Averages

Calculate the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_averages <- pg_crime |> mutate(average_incidents = (Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec) / 12) |> mutate(total_incidents = (Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec))
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

ANSWER HERE 
There are an astounding amount of what I believe are car accidents every month. Car accidents are the crime with the highest average amount of incidents per month. Homicides are the crime with the lowest amount of crimes per month. Now I think it's generally acknowledged that homicides should be near the bottom of the crimes per month, but I think it could be worth comparing that number to state averages. I'll round that homicides number down to 5 to be generous, but that means that in a year at least 60 people are getting murdered. That number seems like it could be very high to me. Car accidents also seems worth checking out to me? That number also seems like it could be really high. What are the police reporting as a car accident? Are these serious accidents or does it include fender benders and scrapes? Is almost 500 accidents a year normal statewide or nationally?

Disclaimer:
I was not able to figure out (though I think I could if I tried a little harder) how to get the N/A inputs to not mess up the residential robbery data (I think it has something to do with na.omit or na.exclude but I couldn't get it to work), so the averages data set and all the following data sets will not include that data (In real life I would jsut hand calculate it).

### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages |> mutate(june_change = (Jun - average_incidents) / average_incidents) |> mutate(july_change = (Jul - average_incidents) / average_incidents) |> mutate(august_change = (Aug - average_incidents) / average_incidents)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

ANSWER HERE I would say that it mostly disproves the theory, or at least proves that the theory is too broad in scope. Yes some crime does look it increases slightly in some months. Homicide rates do appear to go up across the board, but across all those months it looks like theft decreased slightly. So to that 'all crime" increases in summer months I think would be wrong, but looking at the data over many years to see if certain crimes do increase in the summer could be worth looking into.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE There are nearly six homicides a month in Prince Geroges County, according to police reports.

Data from the year end crime report shows that there were 70 homicides in Prince George's County in 2024. On average close to six people were killed per month, with the most killings happening in June.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
