---
title: "HW7_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```



### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
m_and_m_data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10. Be sure to use set.seed(1234) before you start.
```{r}
set.seed(1234)

sample_5 <- m_and_m_data |> 
  sample_n(5)
sample_10 <- m_and_m_data |> 
  sample_n(10)
```




### Task 4: Calculate the mean and standard deviation of the two samples, using one of the colors.
```{r}
sample_5 |> summarize(mean5 = mean(red), sd5 = sd(red))
sample_10 |> summarize(mean10 = mean(red), sd10 = sd(red))
```



### Task 5: Calculate the mean and standard deviation of the original dataset, using the same color as Task 4
```{r}
m_and_m_data |> summarise(mean = mean(red), sd = sd(red))
```



### Task 6: Make three histograms: one each for sample5, sample10 and the original dataset. Use a consistent binwidth.
```{r}
sample_5 |> 
  ggplot() + 
  geom_histogram(aes(x = red), binwidth = 5) +
  geom_vline(aes(xintercept = mean(red)), color = "red", linetype = "dashed", size = 1) + 
   geom_vline(aes(xintercept = mean(red) - sd(red)), color = "blue", linetype = "dashed", size = 1) + 
   geom_vline(aes(xintercept = mean(red) + sd(red)), color = "blue", linetype = "dashed", size = 1) 

sample_10 |> 
  ggplot() + 
  geom_histogram(aes(x = red), binwidth = 5) +
  geom_vline(aes(xintercept = mean(red)), color = "red", linetype = "dashed", size = 1) + 
   geom_vline(aes(xintercept = mean(red) - sd(red)), color = "blue", linetype = "dashed", size = 1) + 
   geom_vline(aes(xintercept = mean(red) + sd(red)), color = "blue", linetype = "dashed", size = 1) 

m_and_m_data |> 
  ggplot() + 
  geom_histogram(aes(x = red), binwidth = 5) +
  geom_vline(aes(xintercept = mean(red)), color = "red", linetype = "dashed", size = 1) + 
   geom_vline(aes(xintercept = mean(red) - sd(red)), color = "blue", linetype = "dashed", size = 1) + 
   geom_vline(aes(xintercept = mean(red) + sd(red)), color = "blue", linetype = "dashed", size = 1) 
```



### Task 7: Answer some questions

Compare the three sets of means, standard deviations and histograms. Which are more helpful to you in understanding the accuracy of the samples?

The histograms and the difference in readability between the samples and the total population really highlight the need for an accurate sample. The mean and the standard devaition alone make it hard to contextualize the data.

What does this exercise tell you about the relationship between sample size and estimate accuracy?

The larger the sample size the closer the estimate accuracy will get.

How does random sampling introduce variation into statistical estimates?

When you ssmple randomly the data you work with will theoretically be drawn from different individuals every time which will vary the results.

If you repeated this process, would the samples always be identical? Why or why not?

No because I would want to have my samples be more varied. Using one sample all the time risks messing up my data if that sample is not representative of what is really going on, if that makes sense.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
